---
title: "STAT 231: Problem Set 1B"
author: "Willie Cohen"
date: "due by 5 PM on Friday, February 26"
output: pdf_document
---

Series B homework assignments are designed to help you further ingest and practice the material covered in class over the past week(s).  You are encouraged to work with other students, but all code must be written by you and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps1B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps1B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
# Load packages here 
library(mdsr)   
library(tidyverse)
library(nasaweather)


knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",   # slightly smaller font for code
  message=FALSE)   
```

\newpage 

# If you discussed this assignment with any of your peers, please list who here:

> ANSWER: Amaaya Smole



\newpage
# MDSR Exercise 2.5 (modified)

Consider the data graphic for Career Paths at Williams College at: https://web.williams.edu/Mathematics/devadoss/careerpath.html.  Focus on the graphic under the "Major-Career" tab.

a. What story does the data graphic tell?  What is the main message that you take away from it?

> ANSWER: The first main message I take away from the data graphic is that the college sends people from a variety of different majors to all sorts of different careers. I would say it essentially implies that in going to Williams you can end up in any career path. Further I think it illustrates that certain more popular majors, like economics, history, and poltical science spread people out to many differnet careers while more focused majors, like Biology and Chemistry have more streamlined outcomes for students.  

b. Can the data graphic be described in terms of the taxonomy presented in this chapter?  If so, list the visual cues, coordinate system, and scale(s).  If not, describe the feature of this data graphic that lies outside of that taxonomy.

> ANSWER: The data graphic uses many visual cues including color to represent different majors as a categorical variable, area of the lines between the majors and the positions (i.e larger area meant more students went into that field), position in that majors that are more alike are closer to eachother and also positioned closer to the positions that they are more likely to go into (i.e biology and chemistry are adjacent and near the health and medicine field).

> In terms of cooridnate systems, this one cannot really be defined with the books taxonomy, as though the data is presented in a circle it is not exactly a polar coordinate system as there is not really a coordinate grid made up from this circle. The circle is just employed in order display all of the majors and job fields in one place. The position of the categories around the circles does seem to have some meaning, meaning the angle coordinate in a polar system would have some meaning, but again its not exactly a by-the-book example of a polar system as the radial coordinates don't have a real meaning. 

> The figure uses a categorical scale, sorting the data points by majors and fields of work instead of numbers or time. 

c. Critique and/or praise the visualization choices made by the designer.  Do they work? Are they misleading?  Thought-provoking?  Brilliant?  Are there things that you would have done differently?  Justify your response.

> ANSWER: The visualization choices made by the designer are extremely creative and appear to representative of the dataset they are working with. Though there is some chaos when all of the lines drawn between majors and jobs are viewed together, I think this sort of chaos is representative of the intended message of the graph, being that Williams graduates are well equipped to enter a variety of fields after graduation. Further, the use of color and the varying areas of the lines to show how many people took that path are very clear and well done. Also, the ability for the graphic to be viewed in both directions (i.e major -> job and job -> major) allows one to observe relationships in both directions which I found very creative. If there was one thing to change, the orientation of the major and job field titles are tough to read and rather small since they are not all upward facing, but that is somewhat of a small problem in the context of the graphic. 


\newpage
# Spot the Error (non-textbook problem)

Explain why the following command does not color the data points blue, then write down the command that will turn the points blue.

> ANSWER: In the original command we are defining our plot object with geom_point, so when we are trying to set the color it is being set baed upon each of the data points value for the "blue" variable which is not a part of the dataset. Then, since all of the values for each data point under our imaginary "blue" are identical (as they do not exist) it sets each point to the same default color as it would if we were coloring based upon the values of a categorical variable. In fact, you could write any word in the parentheses containing "blue" and we would see the exact same plot with the red colored points.  In the correct plot, we define our plot under the main ggplot function, which then allows us to custmoize the color of the points specifically using the geom_point function.

```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(color = "blue")

```


\newpage
# MDSR Exercise 3.6 (modified)

Use the `MLB_teams` data in the `mdsr` package to create an informative data graphic that illustrates the relationship between winning percentage and payroll in context. What story does your graph tell?

> ANSWER: In order to best illustrate the desired relationship between winning percentage and payroll in context, I defined an overall plot with all teams winning percentages and payrolls (meaning all teams will be listed once for each year) across all years of the data set as well as faceted plots for winning percentages and payroll on a year by year basis. I colored each point based upon what league, either AL or NL, the team plays in as differences between the two leagues are often important when considering league-wide metrics or statsitcs. My graphic shows a moderate postive relationship is visible between payroll and winning percentage which does not vary much by league. Further, I think my data may illustrate possible diminishing marginal returns to payroll when it comes to regualr season winning percentage, something visible in both the main and faceted plots. The highest spending teams tended to be among the winningest in most years, but those winning percentages were almost always matched or surpassed by at least one lower spending team in each season, seen in the faceted plots. These lower spending teams tended to be in the middle of the pack among spending, which again indicates that there is some positive correlation. Then, in viewing the main plot, we see there is almost an even split between the winningest teams across the data set from highest speding and moderate spending teams, with lower spending teams more often ending up at the bottom. 

```{r}
g <-ggplot(data = MLB_teams, aes(x = payroll, y = WPct, color = lgID)) + 
  xlab("Payroll ($)") + ylab("Winning Percentage") +
  ggtitle("Relationship Between Team Payroll and Regular Season Winning Percentage \n2008-2014") 
  
g + geom_smooth(se = FALSE) + geom_point()
g + facet_wrap(~yearID) + geom_point()

```


\newpage
# MDSR Exercise 3.10 (modified)

Using data from the `nasaweather` package, use the `geom_path()` function to plot the path of each tropical storm in the `storms` data table (use variables `lat` (y-axis!) and `long` (x-axis!)).  Use color to distinguish the storms from one another, and use facetting to plot each `year` in its own panel.  Remove the legend of storm names/colors by adding `scale_color_discrete(guide="none")`. 

Note: be sure you load the `nasaweather` package and use the `storms` dataset from that package!

```{r}
library(nasaweather)
stormsdata <- nasaweather::storms
g <-ggplot(data = stormsdata, aes(x=long, y = lat, color = name)) + geom_path() + 
  scale_color_discrete(guide = "none")
g + facet_wrap(~year) +
  xlab("Longitude") + ylab("Latitude") + 
  ggtitle("Path of Tropical Storms by Year (1995-2000)")
```


\newpage
# Calendar assignment check-in

For the calendar assignment:

- Identify what questions you are planning to focus on
- Describe two visualizations (type of plot, coordinates, visual cues, etc.) you imagine creating that help address your questions of interest 
- Describe one table (what will the rows be?  what will the columns be?) you imagine creating that helps address your questions of interest

Note that you are not wed to the ideas you record here.  The visualizations and table can change before your final submission.  But, I want to make sure your plan aligns with your questions and that you're on the right track.

> ANSWER: Some questions I would like to focus on are as follows:

> 1) What time do I start (school) work each day and what time do I finish (school) work at night?
> 2) How much time do I sleep and how does it vary throughout the days of the week (also when do I go to sleep/wake up)?
> 3) How many hours a day do I spend in my room/outside/in other buildings on campus?

> GRAPHICS: For the second question, regarding time slept, I could created a faceted line graph/time series by week with an accompanying bar graph to show average hours slept by day of the week.
> For the third question I could try to build a sort of geographical plot (like ones we've seen with maps of the US) but with a map of the Amherst campus and then color the buildings I spend my time in accordingly.

> For the table I could track what time I start working on schoolwork and what time I finish working on it each night. So, the rows would be dates, and the columns would be time started, time finished, and time worked. 


 